Classification of shapes

### 2019-06-28 manuscript/Manuscript2.Rnw
Manuscript Update.

\subsection{Curve Completion1}
Out goal is to complete $[\C]_{obs}$ or ($[\C^{\star}]_{obs}$) by randomly drawing from the distribution on the missing part of the shape conditional on the observed part.  Specifically, for a fixed translation, rotation, scaling, and parameterization, we seek to approximate the distribution of $\C_{mis}|\C_{obs}$ and use random draws from this distribution to complete to curve.  Here this distribution will be approximated non-parametrically and hot-deck type multiple imputation \cite{AndridgeLittle2010} will be used similar in spirit to the idea of predictive mean matching (PMM) (\cite{PMM}).  The basic idea of PMM is to to take partially observed records (termed the donees) and find fully observed records that are close to a particular donee (termed donors).  Random draws from the set of donors are then used to complete the partially observed donees.

The same type of procedure will be performed here, but instead of traditional statistical records, here shapes are considered.  The spirit of the procedure proposed here is similar to PMM in that first the partially observed shape (the donee) is measured against fully observed shapes to find potential donors.  Once the set of potential donors is defined, one of these shapes is randomly drawn and used to complete the partially observed donee.

The observed region $\mathcal{R}_j$ associated with a partially observed curve $\C^p_j$ is the subinterval $[0,t_j]$ with $t_j<1$ for all $j=1,\ldots,n_p$. Suppose that $\C^p_j(0)=\mathbf{a}_j:=(a_{1j},a_{2j})^T$ and $\C^p_j(t_j)=\mathbf{b}_j:=(b_{1j},b_{2j})^T$. Then the set of curves that can possibly comprise the missing segment of curve $\C^p_j$ is 
$$
\mathcal{X}_j:=\{\X: [t_j,1] \to \Rtwo: \X(t_j)=\mathbf{b}, \X(1)=\mathbf{a}\} .
$$

@@

%I also want to minimize over subsets of the unit circle.
%With no scaling
%add the p super script because these are measuring distances between partial shapes and parts of the complete shape.  
$$
\Phi^{p,\star}_{\theta_{ij}}(C_j^p,C_i^f) = d^{\star}(\C_j^p, \C_i^f|_{s}) = \inf_{(s,\gamma,O ) \in \uc \times \Gamma_I\times SO(2)} \|Q_{C_j^p , OQ_{C_i^f}(\gamma)|_{s}\sqrt{\gamma'}\|_{\ltwo}^2
$$

@@

Now for a fixed $j$, after computing $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)$ for all $i = 1, 2, \cdots, n_f$, the indices of the donor set are defined to be $D^{\star}_j = \{i : \Phi^{\star}_{\theta_j}(C_j^p,C_i^f) \le \Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(k)} \}$.  $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(k)}$ is the $k$-th smallest value of $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(k)}$ over all $i = 1, 2, \cdots, n_f$ (e.g. $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(1)}$ and $\Phi^{\star}_{\theta_ij}}(C_j^p,C_i^f)_{(n_p)}$ are the minimum and the maximum, respectively).

@@

%How do I incorporate s here?  This is different than what Karthik had written.  He was minimizing across everything.  We first do partial matching and then once that is done we solve for X by minimizing distances between closed curves.  So no transformations need to be found.  Only find X.
$$
X_j^{\star *} =\argmin_{X_j \in \mathcal{X}_j} \|Q_{C_j^p \circ X_j}, O^*_{ij}Q_{C_i^f}(\gamma^*_{ij})\sqrt{\gamma^*_{ij}'}\|_{\ltwo}^2
$$

@@

$$
X_{ij}^{*} =\argmin_{X_{ij} \in \mathcal{X}_{ij}} \|Q_{C_j^p \circ X_{ij}}, \sigma^* O^*_{ij}Q_{C_i^f}(\gamma^*_{ij})\sqrt{\gamma^*_{ij}'}\|_{\ltwo}^2
$$.

@@

\subsection{Partial matching}

@@

\subsection{Curve completion}
The observed region $\mathcal{R}_j$ associated with a partially observed curve $\C^m_j$ is the subinterval $[0,t_j]$ with $t_j<1$ for all $j=1,\ldots,N$. Suppose that $\C^m_j(0)=\mathbf{a}_j:=(a_{1j},a_{2j})^T$ and $\C^m_j(t_j)=\mathbf{b}_j:=(b_{1j},b_{2j})^T$. Then the set of curves comprising the missing segment of curve $\C^m_j$ is

$$
\mathcal{X}_j:=\{\X: [t_j,1] \to \Rtwo: \X(t_j)=\mathbf{b}, \X(1)=\mathbf{a}\} 
$$.

For a partially observed $\C_j^m:[0,t_j]$ and an $X \in \mathcal{X}_j$ with $j=1,\ldots,n$, define its completion to be the concatenated closed curve

$$
C_j \circ X(t):= \C_j^m(t) \mathbb{I}_{t \in [0,t_j]}+\X(t) \mathbb{I}_{(t_j,1]}.
$$

Denote by $Q_{\C^m_j \circ X}$ the square-root transform of $\C^m_j \circ X$. For each $j=1,\ldots,n$, let $\Theta_j:=\uc \times \Gamma_I \times SO(2) \times \mathcal{X}_j$, and recall that $\mathcal{C}^o$ and $\mathcal{C}^m$ denote the sets of fully and partially observed curves, respectively. For a fixed $C \in \mathcal{C}$, for each $j=1,\ldots,n$ define the cost functional $\Phi_{\theta_j}:\mathcal{C}^m \times \mathcal{C} \to \mathbb{R}$ by
\begin{align*}
\Phi_{\theta_j}(C_j^m,C)&:=d^2(C_j^m \circ X_j, OC(\gamma)), \quad \theta_j \in \Theta_j\\
&=\inf_{(s,\gamma,O) \in \uc \times \Gamma_I\times SO(2)}\|Q_{C_j^m \circ X_j}, OQ_{C}(\gamma)\sqrt{\gamma'}\|_{\ltwo}^2.
\end{align*}
The optimal shape completion of a partially observed curve $Q_{C_j^m}, j=1,\ldots,n$ is $Q_{C_j^m\circ X_j^*}$, where $X_j^*$ is obtained from the solution set of:
\begin{equation}
\label{opt2}
\theta^*_j:=(s_j^*,\gamma_j^*,O_j^*,X_j^*)=\argmin_{\theta_j \in \Theta_j} \Phi_{\theta_j}(C_j^m,C).
\end{equation}
In the expression above $s_j^*$ corresponds to the optimal point at which $\uc$ was unwrapped in order to identify $\Gamma$ with $\uc \times \Gamma_I$; $\gamma_j^*:[0,1] \to [0,1]$ and $O_j^*$ represents the optimal reparameterization of the curve $C^m_j \circ X^*_j$. The use of a valid distance on the quotient shape space $\mathcal{Q}_s$ allows us to apply the shape transformations on $Q_{C_i}, i=1\ldots,n$ in the variational problem with introducing any arbitrariness. The (product) group structure of $\uc \times \Gamma_I\times SO(2)$, and its action on $\mathcal{Q}$, ensures that if the transformations were to have been applied to $Q_{C^m_j \circ X^*}$, the resulting solution set would instead contain the corresponding group inverses. 
(INSERT ILLUSTRATIVE FIGURE).

>>>>

\subsection{Curve Completion1}
Out goal is to complete $[\C]_{obs}$ or ($[\C^{\star}]_{obs}$) by randomly drawing from the distribution on the missing part of the shape conditional on the observed part.  Specifically, for a fixed translation, rotation, scaling, and parameterization, we seek to approximate the distribution of $\C_{mis}|\C_{obs}$ and use random draws from this distribution to complete to curve.  Here this distribution will be approximated non-parametrically and hot-deck type multiple imputation \cite{AndridgeLittle2010} will be used similar in spirit to the idea of predictive mean matching (PMM) (\cite{PMM}).  The basic idea of PMM is to to take partially observed records (termed the donees) and find fully observed records that are close to a particular donee (termed donors).  Random draws from the set of donors are then used to complete the partially observed donees.  This is repeated $M$ times to create $M$ completed data sets.

The same type of procedure will be performed here, but instead of traditional statistical records, here shapes are considered.  The spirit of the procedure proposed here is similar to PMM in that first the partially observed shape (the donee) is measured against fully observed shapes to find potential donors.  Once the set of potential donors is defined, one of these shapes is randomly drawn and used to complete the partially observed donee.  This process is crepeated $M$ times to create $M$ completed shapes.

The observed region $\mathcal{R}_j$ associated with a partially observed curve $\C^p_j$ is the subinterval $[0,t_j]$ with $t_j<1$ for all $j=1,\ldots,n_p$. Suppose that $\C^p_j(0)=\mathbf{a}_j:=(a_{1j},a_{2j})^T$ and $\C^p_j(t_j)=\mathbf{b}_j:=(b_{1j},b_{2j})^T$.Then the set of curves that can possibly comprise the missing segment of curve $\C^p_j$ is

$$
\mathcal{X}_j:=\{\X: [t_j,1] \to \Rtwo: \X(t_j)=\mathbf{b}, \X(1)=\mathbf{a}\} 
$$
.

@@

%I also want to minimize over subsets of the unit circle.
%With no scaling
%add the p super script because these are measuring distances between partial shapes and parts of the complete shape.  
$$
\Phi^{p,\star}_{\theta_{ij}}(C_j^p,C_i^f) = d^{\star}(\C_j^p, \C_i^f|_{s}) = \inf_{(s,\gamma,O ) \in \uc \times \Gamma_I\times SO(2)} \|Q_{C_j^p} , OQ_{C_i^f}(\gamma)|_{s}\sqrt{\gamma'}\|_{\ltwo}^2
$$

@@

Now for a fixed $j$, after computing $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)$ for all $i = 1, 2, \ldots, n_f$, the indices of the donor set are defined to be $D^{\star}_j = \{i : \Phi^{\star}_{\theta_j}(C_j^p,C_i^f) \le \Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(k)} \}$. $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(k)}$ is the $k$-th smallest value of $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(k)}$ over all $i = 1, 2, \ldots, n_f$ (e.g. $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(1)}$ and $\Phi^{\star}_{\theta_{ij}}(C_j^p,C_i^f)_{(n_p)}$ are the minimum and the maximum, respectively).

@@

%Something is wrong here latex wise
$$
X_j^{\star *} =\argmin_{X_j \in \mathcal{X}_j} \|Q_{C_j^p \circ X_j}, O^{*}_{ij}Q_{C_i^f} (\gamma^{*}_{ij})\sqrt{\gamma'^{*}_{ij}}\|_{\ltwo}^2
$$

%How do I incorporate s here?  This is different than what Karthik had written.  He was minimizing across everything.  We first do partial matching and then once that is done we solve for X by minimizing distances between closed curves.  So no transformations need to be found.  Only find X.

@@

$$
X_j^{*} =\argmin_{X_j \in \mathcal{X}_j} \|Q_{C_j^p \circ X_j}, O^{*}_{ij}Q_{C_i^f} (\gamma^{*}_{ij})\sqrt{\gamma'^{*}_{ij}}\|_{\ltwo}^2
$$.

@@

\subsection{Classification}

@@

\section{Simulation Study}
This simulation study is based on the shapes of the occlusal surfaces of of extant bovid molars both upper and lower.  Specifically, teeth from the following seven tribes: Alcelaphini, Antilopini, Bovini, Hippotragini, Neotragini, Reduncini, and Tragelaphini.  Specific counts of the sample sizes of each tooth type and tribe can be found in table \ref{sampsize}.

For each tooth, the raw data was made up of 60 points traced around the occlusal surface of the tooth.  For each tooth, these 60 points were split into two groups roughly divided by a line connecting the mesostyle to the entostyle/metastylid to the ectostylid of mandibular teeth/upper teeth.
An example of this can be seen in Figure \ref{ToothCutter}.  This cut was chosen for the simulation study and this is a break point commonly observed by in bovid fossils by biological anthropologists.

For each half of a tooth, the tooth was matched against all the complete teeth in the tooth type to find $k$ suitable donor teeth, the tooth was then completed $M$ times by randomly drawing $M$ teeth with replacement from the set of $k$ donor teeth, and using the donor tooth to complete the tooth as described in section XXX.

tooth <- "LM1"

complete_shape <- t(ptsTrainList[[1]][[1]])
plot(t(complete_shape), type = "l")
points(t(complete_shape), pch = 16)

side <- 1
partial_shape <- t(tooth_cutter(ptsTrainList[[tooth]][[1]])[[side]])
plot(t(complete_shape), type = "l")
points(t(complete_shape), pch = 16)
points(t(partial_shape), pch = 16, col = "red")
points(t(partial_shape), type = "l", col = "red")

side <- 2
partial_shape <- t(tooth_cutter(ptsTrainList[[tooth]][[1]])[[side]])
points(t(partial_shape), pch = 16, col = "blue")
points(t(partial_shape), type = "l", col = "blue")
@

\caption{Example cuts for an Alcelaphini LM1 (DSCN2871)}
\label{ToothCutter}
\end{figure}

<<>>=
  k = 5
  M = 5
  side <- 1
  tooth <- "LM3"
load(paste0("/Users/gregorymatthews/Dropbox/shapeanalysisgit/results/results20190610_side=",side,"_k=",k,"_M=",M,"_tooth=",tooth,".RData"))

plot(t(results_list[[1]]$imputed_partial$imputed[[1]]),col = "red", type = "l", xlim = c(-200,700), ylim = c(-200,200))
points(t(results_list[[1]]$imputed_partial$partial_obs))
points(t(results_list[[1]]$imputed_partial$imputed[[2]]),col = "orange", type = "l")
points(t(results_list[[1]]$imputed_partial$imputed[[3]]),col = "gold", type = "l")
points(t(results_list[[1]]$imputed_partial$imputed[[4]]),col = "darkgreen", type = "l")
points(t(results_list[[1]]$imputed_partial$imputed[[5]]),col = "blue", type = "l")

greg <- c()
for (i in 1:length(results_list))
  {
  greg[i] <- results_list[[i]]$truth$Tribe
  }
  table(greg)
  sum(table(greg))
@

  \begin{table}[h!]
  \begin{center}
\begin{tabular}{ c c c c c c c c  }
  & Alcelaphini  & Antilopini & Bovini &Hippotragini & Neotragini & Reduncini & Tragelaphini & Total \\ 
 LM1 & 106 & 27 & 22 & 26 & 45 & 76 & 53 & 355 \\ 
 LM2 & 117 & 30 & 25 & 37 & 55 & 90 & 53 & 407\\ 
 LM3 & 117 & 30 & 19 & 34 & 53 & 96 & 62 & 411 \\ 
 UM1 & 117 & 30 & 23 & 37 & 43 & 80 & 75 & 405 \\ 
 UM2 & 118 & 30 & 23 & 41 & 53 & 97 & 94 & 456\\ 
 UM3 & 71 & 30 & 17 & 58 & 52 & 110 & 78 &  416\\ 
\end{tabular}
\end{center}
\caption{Sample sizes for the simulation study by tooth type and tribe}
\label{sampsize}
\end{table}

%These are the numbers after the shapes that are scale incorreclty have been removed.  
LM1 - 355
LM2 - 407
LM3 - 411
UM1 - 405
UM2 - 456
UM3 - 416

DSCN2601 example

%<<>>=
%  include_graphics("/Users/gregorymatthews/Dropbox/shapeanalysi%sgit/manuscript/fig/DSCN2601.JPG") 
%@
